Nuclei Segmentation in Histopathology images using Deep Neural Networks

### 2016-11-07 Paper.tex
new

\section{Related work}
\label{sec:related}
Image segmentation and object recognition in medical imagining have 
been tackled for many years and many approaches have been found. 
Methods based on mathematical morphological operators have been 
widely used and applied for segmentation and feature extraction, 
however such methods rely on the user to manually define the 
information that is relevant for his task \cite{irshad2014methods}. 
Segmentation methods based on a strong priori information are also 
possible and yield comparable results in terms of accuracy, in 
\cite{ranefall2016fast} they segment the picture using a ellispe fit 
which is considerably faster than other state of the art methods. Graph-
based tools for segmentation are also possible in histopathology 
\cite{ta2009graph}. \cite{manivannanlocal} focus on gland 
segmentation, they achieve good results by predicting image patches as 
a whole, via a classification scheme. The output confidence vector is 
then preprocessed to obtain the final segmentation. Alternative methods 
have arisen from the achieving Convolutional neural networks 
\cite{lecun}, in particular strong incentive rose with the very impressive 
results that they achieved in the pascal voc challenge with 
\cite{ImageNet}. Recent advances in deep neural network, and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition.  Deep neural network have also been 
used for different task and have yielded very impressiv results, we can think of 
semantic segmentation problems, where \cite{long2015fcn} use "de-
convolution layers" and up-sampling in order to identify and precisely 
locate objects within a picture. Different architecture arising from 
different intuition are also possible and have been applied in this paper
Talk about related work.

@@

%% Data augmentation and Hyper parameter training: Training scheme
%% Check number of images and number of transformations
To train our models, as the number of available annotated is scarse, we 
used a great number of transformation for the data augmentation. From 
a original size of 33 annotated images, adding flips, rotations, bluriness 
and random elastic deformations enabled us to have more then 
$400 000$ training images. 
We also try out several hyperparameter configurations: the learning rate and 
momentum for the stochastic gradient descent, the weight decay value.  
In practice, we found that hyperparameters tuning didn't influence the 
scores much, the exepction being the learning rate. If the learning rate 
was not of the right magnitude the given network did not seem to learn. 
We also experienced with different 
initialization value and if possible, we also considered pretrained layers. 
Using pretrained layers made learning more efficient and made score 
values more robust.

\section{Different architectures/Results}
\label{sec:method}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned. The final architecture that we tried is the UNet. This network combines both contribution of the above networks, the UNet has a similar architecture to the DeconvNet with skip layers.
\section{Result}
\label{sec:result}

\begin{tabular}{|c|c|c|c|c|c|}
 & UNet & Baochuan & DeconvNet & FCN & Ensemble \\
%\hrule
Accuracy          &       0.89 &       0.94 & 0.97\\
IU   &       0.70 &       0.76 &     0.86 \\
Recall                     &       0.75 &       0.74  &     0.86 &&\\
Precision                  &       0.64 &       0.74 &     0.86 &&\\
F1                         &       0.69 &       0.74 &     0.86 &&  \\
TP             &       0.75 &       0.74 &     0.86 && \\
TN             &       0.92 &       0.96 &     0.98 &&\\
Performance                &       0.84 &       0.85 &     0.92 &&\\
Pixel error                &       0.11 &       0.06 &     0.03 &&\\
\end{tabular}

\section{Discussion}
\label{sec:discussion}
What's next?

%\section{Formatting your paper}
%\label{sec:format}
%
%All printed material, including text, illustrations, and charts, must be kept
%within a print area of 7 inches (178 mm) wide by 9 inches (229 mm) high. Do
%not write or print anything outside the print area. The top margin must be 1
%inch (25 mm), except for the title page, and the left margin must be 0.75 inch
%(19 mm).  All {\it text} must be in a two-column format. Columns are to be 3.39
%inches (86 mm) wide, with a 0.24 inch (6 mm) space between them. Text must be
%fully justified.
%
%\section{PAGE TITLE SECTION}
%\label{sec:pagestyle}
%
%The paper title (on the first page) should begin 1.38 inches (35 mm) from the
%top edge of the page, centered, completely capitalized, and in Times 14-point,
%boldface type.  The authors' name(s) and affiliation(s) appear below the title
%in capital and lower case letters.  Papers with multiple authors and
%affiliations may require two or more lines for this information.
%
%\section{TYPE-STYLE AND FONTS}
%\label{sec:typestyle}
%
%To achieve the best rendering both in the proceedings and from the CD-ROM, we
%strongly encourage you to use Times-Roman font.  In addition, this will give
%the proceedings a more uniform look.  Use a font that is no smaller than nine
%point type throughout the paper, including figure captions.
%
%In nine point type font, capital letters are 2 mm high.  If you use the
%smallest point size, there should be no more than 3.2 lines/cm (8 lines/inch)
%vertically.  This is a minimum spacing; 2.75 lines/cm (7 lines/inch) will make
%the paper much more readable.  Larger type sizes require correspondingly larger
%vertical spacing.  Please do not double-space your paper.  True-Type 1 fonts
%are preferred.
%
%The first paragraph in each section should not be indented, but all the
%following paragraphs within the section should be indented as these paragraphs
%demonstrate.
%
%\section{MAJOR HEADINGS}
%\label{sec:majhead}
%
%Major headings, for example, "1. Introduction", should appear in all capital
%letters, bold face if possible, centered in the column, with one blank line
%before, and one blank line after. Use a period (".") after the heading number,
%not a colon.
%
%\subsection{Subheadings}
%\label{ssec:subhead}
%
%Subheadings should appear in lower case (initial word capitalized) in
%boldface.  They should start at the left margin on a separate line.
% 
%\subsubsection{Sub-subheadings}
%\label{sssec:subsubhead}
%
%Sub-subheadings, as in this paragraph, are discouraged. However, if you
%must use them, they should appear in lower case (initial word
%capitalized) and start at the left margin on a separate line, with paragraph
%text beginning on the following line.  They should be in italics.
%
%\section{PRINTING YOUR PAPER}
%\label{sec:print}
%
%Print your properly formatted text on high-quality, 8.5 x 11-inch white printer
%paper. A4 paper is also acceptable, but please leave the extra 0.5 inch (12 mm)
%empty at the BOTTOM of the page and follow the top and left margins as
%specified.  If the last page of your paper is only partially filled, arrange
%the columns so that they are evenly balanced if possible, rather than having
%one long column.
%
%In \LaTeX, to start a new column (but not a new page) and help balance the
%last-page column lengths, you can use the command ``$\backslash$pagebreak'' as
%demonstrated on this page (see the \LaTeX\ source below).
%
%\section{PAGE NUMBERING}
%\label{sec:page}
%
%Please do {\bf not} paginate your paper.  Page numbers, session numbers, and
%conference identification will be inserted when the paper is included in the
%proceedings.
%
%\section{ILLUSTRATIONS, GRAPHS, AND PHOTOGRAPHS}
%\label{sec:illust}
%
%Illustrations must appear within the designated margins.  They may span the two
%columns.  If possible, position illustrations at the top of columns, rather
%than in the middle or at the bottom.  Caption and number every illustration.
%All halftone illustrations must be clear black and white prints.  If you use
%color, make sure that the color figures are clear when printed on a black-only
%printer.
%
%Since there are many ways, often incompatible, of including images (e.g., with
%experimental results) in a \LaTeX\ document, below is an example of how to do
%this \cite{Lamp86}.

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{RGB_1}}
%  \vspace{2.0cm}
  \centerline{Patient 498959}\medskip
\end{minipage}
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

@@

%\section{FOOTNOTES}
%\label{sec:foot}
%
%Use footnotes sparingly (or not at all!) and place them at the bottom of the
%column on the page on which they are referenced. Use Times 9-point type,
%single-spaced. To help your readers, avoid using footnotes altogether and
%include necessary peripheral observations in the text (within parentheses, if
%you prefer, as in this sentence).
%

%\section{COPYRIGHT FORMS}
%\label{sec:copyright}
%
%You must include your fully completed, signed IEEE copyright release form when
%you submit your paper. We {\bf must} have this form before your paper can be
%published in the proceedings.  The copyright form is available as a Word file,
%a PDF file, and an HTML file. You can also use the form sent with your author
%kit.

@@

%List and number all bibliographical references at the end of the
%paper.  The references can be numbered in alphabetic order or in
%order of appearance in the document.  When referring to them in the
%text, type the corresponding reference number in square brackets as
%shown at the end of this sentence \cite{C2}.

>>>>

\section{Related work}
\label{sec:related}
Image segmentation and object recognition in medical imagining have 
been tackled for many years and many approaches have been found. 
Methods based on mathematical morphological operators have been 
widely used and applied for segmentation and feature extraction, 
however such methods rely on the user to manually define the 
information that is relevant for his task \cite{irshad2014methods}. 
Segmentation methods based on a strong priori information are also 
possible and yield comparable results in terms of accuracy, in 
\cite{ranefall2016fast} they segment the picture using a ellispe fit 
which is considerably faster than other state of the art methods. Graph-
based tools for segmentation are also possible in histopathology 
\cite{ta2009graph}. \cite{manivannanlocal} focus on gland 
segmentation, they achieve good results by predicting image patches as 
a whole, via a classification scheme. The output confidence vector is 
then preprocessed to obtain the final segmentation. Alternative methods 
have arisen from the achieving Convolutional neural networks 
\cite{lecun}, in particular strong incentive rose with the very impressive 
results that they achieved in the pascal voc challenge with 
\cite{ImageNet}. Recent advances in deep neural network, and in 
particular in their optimization have made them become the state-of-the-
art model for object recognition.  Deep neural network have also been 
used for different task and have yielded very impressiv results, we can think of 
semantic segmentation problems, where \cite{long2015fcn} use "de-
convolution layers" and up-sampling in order to identify and precisely 
locate objects within a picture. Different architecture arising from 
different intuition are also possible and have been applied in this paper.

Talk about related work.

@@

%% Data augmentation and Hyper parameter training: Training scheme
%% Check number of images and number of transformations
To train our models, as the number of available annotated is scarse, we 
used a great number of transformation for the data augmentation. From 
a original size of 33 annotated images, adding flips, rotations, bluriness 
and random elastic deformations enabled us to have more then 
$400 000$ training images. 
We also try out several hyperparameter configurations: the learning rate and 
momentum for the stochastic gradient descent, the weight decay value.  
In practice, we found that hyperparameters tuning didn't influence the 
scores much, the exepction being the learning rate. If the learning rate 
was not of the right magnitude the given network did not seem to learn. 
We therefore fixed the momentum to be 0.9 and the weight decay to be 
$5.10^{-5}$ for all of the experiences, the learning rate was tunned 
according to the model choosen.
We also experienced with different 
initialization value and if possible, we also considered pretrained layers. 
Using pretrained layers made learning more efficient and made score 
values more robust.

\section{Different architectures/Results}
\label{sec:method}
We experience with 4 known arhitectures in semantic segmentation, one 
will be named BaochuanNet, Fully Convolutionnal Net (FCN), DeconvNet, 
UNet. The most basic architecture, BaochuanNet, consists of 4 
convolutionnal layer where each convolutionnal layer has 8 feature map. 
This net, being not deep, has the advantage of being less 
computationnally intensive. FCN is a first attempt of applying "deep 
feature" representations to the task of semantic segmentation. This 
archicture has the advantage of re-using a classical deep learning 
architecture with added on upsampling layer and skip layers. The 
upsampling layers enable the network to learn a pixel level classification 
and the skip layers enable the network to fuse different levels of 
abstraction to the final prediction. This model can be fine tuned with a 
set of pretrained-weights extracted from the classical deep learning 
architecture, for a more detailed explaination please refer to \cite{long2015fcn}. DeconvNet is also based on a classical architecture, 
however, in this network their are no skip layer as one intends to learn 
the proper upsampling threw repeated deconvolution and convolution 
layers. This model can also be fine tuned, for more information please 
refer to the original paper \cite{noh2015learning}. We did not
ensemble classifier as suggested in 
\cite{noh2015learning} as our FCN and DeconvNet classifier achieve very similar scores.
% The final architecture that we tried is the UNet. This network combines 
%both contribution of the above networks, the UNet has a similar 
%architecture to the DeconvNet with skip layers.
\section{Result and discussion}
\label{sec:result}
We conducted our experiments on GPUs via the caffe framework, \cite{caffe}. We present our results in table \ref{tab:resul} where we averaged the metrics other all left out patients. We also added, for illustration purposes \ref{fig:prediction}, image prediction and associated probability maps to evaluate some differences between the classifiers. We notice that the simpler net, Baochuan Net seems to have learnt simple rules based on color information. As we can not notice by the "holes" in the probability map when the interior of the cell is brighter. On the contrary, deeper nets as FCN and DeconvNet have learnt deeper features and seem capable of recognizing whole cells. The small noticable differences in \ref{fig:prediction} are due to a random crop, this results in a slightly shiffted image for each one of the predictions.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline
  & Baochuan & DeconvNet & FCN \\
 \hline
Accuracy  &       0.936 & 0.968 &0.958  \\
IU   &    0.759 &     0.856 & 0.857 \\
Recall     &       0.744  &     0.856 & 0.855 \\
Precision   &       0.741 &     0.858 & 0.878\\
F1    &       0.742 &     0.856 & 0.866 \\
TP       &       0.744 &     0.856 & 0.855  \\
TN      &       0.963 &     0.981 &  0.977\\
Performance    &       0.853 &     0.918 & 0.916\\
Pixel error  &  0.063 &     0.032 & 0.042\\
\hline
\end{tabular}
\caption{\label{tab:res}Results}
\label{tab:res}
\end{table}

%%with Unet
%\begin{tabular}{|c|c|c|c|c|c|}
% & UNet & Baochuan & DeconvNet & FCN & Ensemble \\
%\hrule
%Accuracy &       0.89 &       0.936 & 0.968 &0.958 & \\
%IU   &       0.70 &       0.759 &     0.856 & 0.857 &\\
%Recall    &       0.75 &       0.744  &     0.856 & 0.855 &\\
%Precision    &       0.64 &       0.741 &     0.858 & 0.878&\\
%F1       &       0.69 &       0.742 &     0.856 & 0.866&  \\
%TP             &       0.75 &       0.744 &     0.856 & 0.855 & \\
%TN             &       0.92 &       0.963 &     0.981 &  0.977&\\
%Performance    &       0.84 &       0.853 &     0.918 & 0.916&\\
%Pixel error  &       0.11 &       0.063 &     0.032 & 0.042&\\
%\end{tabular}

\section{PAGE TITLE SECTION}
\label{sec:pagestyle}

The paper title (on the first page) should begin 1.38 inches (35 mm) from the
top edge of the page, centered, completely capitalized, and in Times 14-point,
boldface type.  The authors' name(s) and affiliation(s) appear below the title
in capital and lower case letters.  Papers with multiple authors and
affiliations may require two or more lines for this information.

\section{TYPE-STYLE AND FONTS}
\label{sec:typestyle}

The first paragraph in each section should not be indented, but all the
following paragraphs within the section should be indented as these paragraphs
demonstrate.

% Below is an example of how to insert images. Delete the ``\vspace'' line,
% uncomment the preceding line ``\centerline...'' and replace ``imageX.ps''
% with a suitable PostScript file name.
% -------------------------------------------------------------------------
\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_1}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 498959}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_1}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_2}}
%  \vspace{1.5cm}
  \centerline{Sample from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_2}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{RGB_3}}
%  \vspace{1.5cm}
  \centerline{Another one from patient 581910}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_3}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}
%
\caption{Random annotated samples from the dataset}
\label{fig:annotation}
%
\end{figure}

\begin{figure}[htb]
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{RGB_pred}}
%  \vspace{1.5cm}
  \centerline{Original RGB sample}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[width=4.0cm]{GT_pred}}
%  \vspace{1.5cm}
  \centerline{Associated ground truth}\medskip
\end{minipage}

\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanB}}
%  \vspace{1.5cm}
  \centerline{BaochuanNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{BaochuanP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}

%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetB}}
%  \vspace{1.5cm}
  \centerline{DeconvNet}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{DeconvNetP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
%
\begin{minipage}[b]{.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNB}}
%  \vspace{1.5cm}
  \centerline{FCN}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{0.48\linewidth}
  \centering
  \centerline{\includegraphics[height=4.0cm]{FCNP}}
%  \vspace{1.5cm}
  \centerline{Probability map}\medskip
\end{minipage}
%
\caption{Prediction via different classifiers of a random sample on the left out patient: 572123}
\label{fig:prediction}
%
\end{figure}

@@

\section{FOOTNOTES}
\label{sec:foot}

Use footnotes sparingly (or not at all!) and place them at the bottom of the
column on the page on which they are referenced. Use Times 9-point type,
single-spaced. To help your readers, avoid using footnotes altogether and
include necessary peripheral observations in the text (within parentheses, if
you prefer, as in this sentence).

\section{COPYRIGHT FORMS}
\label{sec:copyright}

You must include your fully completed, signed IEEE copyright release form when
you submit your paper. We {\bf must} have this form before your paper can be
published in the proceedings.  The copyright form is available as a Word file,
a PDF file, and an HTML file. You can also use the form sent with your author
kit.

@@

List and number all bibliographical references at the end of the paper.  The references can be numbered in alphabetic order or in order of appearance in the document.  When referring to them in the text, type the corresponding reference number in square brackets as shown at the end of this sentence \cite{C2}.
